---
title: "Lab 5 TEAM: ALL STARS"
author: "Johnny Tamanaha, Emily Archambault, Eugene Miller, Kaiyuan Fan, Jingzhi Zhang"
date: "9/29/2019"
output: html_document
---
```{r, setup, echo = FALSE, message = FALSE}
library(tidyverse)
milk <- read.csv('state_milk_production.csv')

milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)
```


```{r, echo = FALSE}
year_most_produced <- milk %>%
  group_by(year) %>%
  summarize(total_prod=sum(milk_million)) %>%
  filter(rank(desc(total_prod)) < 2)
```
The year when the most milk was produced in the United States:  
`r year_most_produced[1]`, `r year_most_produced[2]` million lb.

The year when the least milk was produced in the United States:
```{r, echo = FALSE}
year_least<- milk %>%
  group_by(year) %>%
  summarize(total=sum(milk_million)) %>%
  arrange(desc(total))%>%
  top_n(-1)

```
The year when the most milk was produced in the United States:  
`r year_least[1]`, `r year_least[2]` million lb.



```{r, echo = FALSE, message = FALSE, error = FALSE, warning = FALSE}
milk2017 <- milk %>%
  filter(year == 2017)

top5_2017 <- milk2017 %>%
  arrange(desc(milk_produced)) %>%
  top_n(5) %>%
  as.tibble()
```
In 2017, the 5 states that produced the most milk: 

`r top5_2017[1,2]`: `r top5_2017[1,5]` million lb,  
`r top5_2017[2,2]`: `r top5_2017[2,5]` million lb,  
`r top5_2017[3,2]`: `r top5_2017[3,5]` million lb,  
`r top5_2017[4,2]`: `r top5_2017[4,5]` million lb,  
`r top5_2017[5,2]`: `r top5_2017[5,5]` million lb.   

In 2017, the 5 states that produced the least milk:
```{r, echo = FALSE}

```



### Plot For States We Have Lived In

```{r, echo = FALSE, message = FALSE, fig.align='center'}
milkCA <- milk %>%
  filter(state == 'California')

milk_hawaii <- milk %>%
  filter(state=="Hawaii")

milkWA <- milk %>%
  filter(state=="Washington")

milkCT <- milk %>%
  filter(state == "Connecticut")

milkCO <- milk %>%
  filter(state == "Colorado")

milkmean <- milk %>%
  group_by(year) %>%
  summarize('Milk produced in year:' , mean(milk_million))
milkyears <- data.frame(milkmean)
  
ggplot() +
  geom_point(data = milkCA, mapping = aes(x = year, y = milk_million, color = state)) + 
  geom_point(data = milk_hawaii, mapping = aes(x = year, y = milk_million, color = state)) + 
  geom_point(data = milkCT, mapping = aes(x = year, y = milk_million, color = state)) +
  geom_point(data = milkCO, mapping = aes(x = year, y = milk_million, color = state)) +
  geom_point(data = milkWA, mapping = aes(x = year, y = milk_million, color = state)) +
  geom_smooth(data = milkyears, mapping = aes(x = year, y = mean.milk_million., color = "National Average"), se = FALSE) +
  labs(x = "Year", y = "Milk Produced (In Millions of Pounds)", title = "Milk Produced in CA, CT, CO, HI, WA and USA Average", color = "State")
```

Findings: This plot shows that California milk production slowed down after 2007 while nationally milk production continued to increase. The milk production in Hawaii has always been far below the U.S. average, the milk production in Hawaii has had a small decrease over time. Connecticut has similar trends to Hawaii. Both Colorado and Washington show slight upward trends that hold for recent years.


### Individual Sections

### Jingzhi Zhang

Plot of milk produced for year 1998

```{r, echo = FALSE, message = FALSE, fig.align='center'}
milk1998<-milk%>%
  filter(year==1998)%>%
  mutate(milk_million = milk_produced/1000000)
milk1998
ggplot(data = milk1998, mapping = aes(x=milk_produced))+
  geom_density(kernel="gaussian")+
  labs(x="Milk Produced",y="Density",title = "Distribution of The Yearly Milk Production (in millions of pounds) For 1998")+
  theme(text = element_text(color = "blue"))
```
Summary of milk produced by state in 1998:

```{r, echo = FALSE, message = FALSE}
milk1998 <- milk %>%
  filter(year==1998)%>%
  mutate(milk_million = milk_produced/1000000)%>%
  group_by(state) %>%
  summarise(Sum1998=sum(milk_million))
milk1998
```



```{r, echo = FALSE, message = FALSE}
Ave1998 <- milk %>%
  filter(year==1998)%>%
  mutate(milk_million = milk_produced/1000000)%>%
  summarise(Ave1998=mean(milk_million))
Ave1998
```
Average milk produced by each state in 1998 = `r Ave1998` million lb.


```{r, echo = FALSE, message = FALSE}
Me1998 <- milk %>%
  filter(year==1998)%>%
  mutate(milk_million = milk_produced/1000000)%>%
  summarise(Median1998=median(milk_million))
Me1998
```
The median milk produced be each state in 1998 = `r Me1998` million lb.


```{r, echo = FALSE, message = FALSE}
milk1998<-milk%>%
  filter(year==1998)%>%
  mutate(milk_million = milk_produced/1000000)
most<-milk1998%>%
  filter(rank(desc(milk_million))<2)
most
```
the state that produced the most milk in 1998 `r most`.


```{r, echo = FALSE, message = FALSE}
milk1998<-milk%>%
  filter(year==1998)%>%
  mutate(milk_million = milk_produced/1000000)
least<-milk1998%>%
  filter(rank(milk_million)<2)
least
```
the state that produced the least milk in 1998 `r least`.

### Eugene Miller 1999

Plot of milk produced for year 1999:

```{r, echo = FALSE, message = FALSE, fig.align='center'}
milk1999 <- milk %>%
  filter(year == 1999)

ggplot(data = milk1999, mapping = aes(x = milk_million)) +
  geom_density() +
  labs(title = "Distribution of Milk Produced by States for 1999", x = "Milk Produced (in millions of pounds)", y = "Density")

```

Summary of milk produced by state in 1999:
```{r, echo = FALSE, message = FALSE}
milksum1999 <- milk1999 %>%
  group_by(state) %>%
  summarise( milk_million)
milksum1999
```

```{r, echo = FALSE, message = FALSE}
avg_prod <- milk1999$milk_million %>%
  mean(na.rm = TRUE) %>%
 as.tibble()
```
Average state milk production in 1999 = `r avg_prod` million lb.

```{r, echo = FALSE, message = FALSE}
med_prod <- milk1999$milk_million %>%
  median(na.rm = TRUE) %>%
  as.tibble()
```
Median state milk production in 1999 = `r med_prod` million lb.


```{r, echo = FALSE, message = FALSE}
big_prod <- milk1999 %>%
  top_n(1) %>%
  as.tibble()
```
Largest producer of milk in 1999 = `r big_prod[2]` who produced `r big_prod[5]` million lb.

```{r, echo = FALSE, message = FALSE}
sml_prod <- milk1999 %>%
  top_n(-1) %>%
  as.tibble()
```
Smallest producer of milk in 1999 = `r sml_prod[2]` who produced `r sml_prod[5]` million lb.

### Johnny Tamanaha 1999
```{r, fig.align='center'}
milk_1999 <- milk %>%
  filter(year=='1999')
ggplot(milk_1999) +
  geom_density(mapping=aes(milk_million)) +
  labs(title="Distribution of Milk Production in 1999") +
  xlab("Milk Produced (Millions of Pounds)") +
  ylab("Proportion")
```
```{r}
mean <- milk_1999 %>%
  summarize(Mean=mean(milk_million))

median <- milk_1999 %>%
  summarize(Median=median(milk_million))

most_milk <- milk_1999 %>%
  filter(rank(desc(milk_million))<2)

least_milk <- milk_1999 %>%
  filter(rank(milk_million)<2)

mean
```
Mean Milk Production in 1999: `r mean` million pounds
```{r}
median
```
Median Milk Production in 1999: `r median` million pounds
```{r}
most_milk
```
Most Milk Production in 1999: `r most_milk[2]`
```{r}
least_milk
```
Least Milk Production in 1999: `r least_milk[2]`.


### Reflection on What We Found to be Most Helpful:
We found bivariate distributions with a time variable to be the most helpful in visualising this data, as it allowed us to look at how milk production has changed over time state by state. Using specific filters and dyplyr functions we were able to manipulate the data into easily readable formats then were then plotted using ggplot to produce an informative visualization. 

### Reflection on Ethics of This Data Analysis:
This data analysis could hurt milk manufacturers in lesser producing states such as Alaska, or Connecticut because it may divert potential investors towards greater producing states such as Wisconsin and California. In addition, producers of milk alternatives may be happy about this analysis as it shows that in recent years milk production has begun to plateau in California, one of the largest producing states. This could indicate a decreased consumer demand for milk and therefore a movement towards milk alternatives. 

